---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name:_Sharon Gray_


# For 03/17/16

## 4E1
The first line is the likelihood. yi~Normal(mu, sigma)

## 4E2
The posterior distribution includes 2 parameters: the mean (mu) and the standard deviation (sigma).

## 4E3
Pr(mu, sigma | y)= Normal(yi|mu, sigma)Normal(mu|0,10)Uniform(sigma|0,10)/
(integral(Normal(hi|mu, sigma)Normal(mu|0,10)Uniform(sigma|0,10)))d(mu)d(sigma)


## 4M1
```{r}
library(rethinking)
sample_mu <- rnorm( 1e4 , 0 , 10 ) 
sample_sigma <- runif( 1e4 , 0 , 10 ) 
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma ) 
dens( prior_h )
prior_h <- as.data.frame(prior_h)
```
## 4M2
```{r}
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]
flist <- alist(
  height ~ dnorm( mu , sigma ) , 
  mu ~ dnorm( 0 , 10 ) , 
  sigma ~ dunif( 0 , 10 )
)
m4.1 <- map( flist , data=d2 )
precis( m4.1 )
```
# For 03/24/16

##4E4
line 2

##4E5
three parameters: alpha and beta and sigma

## 4M3
yi ~ Normal(mu, sigma)
mui = alpha + beta*xi
alpha ~ Normal(0, 50)
beta ~uniform(0,10)
sigma ~ uniform(0,50)
## 4M4
hi ~ Normal(mui, sigma)
mui = alpha + beta*xi
alpha ~ Normal(147, 50)
beta ~uniform(6,12)
sigma ~ uniform(0,50)

## 4M5
I already am using a positive value of beta to represent growth (height increase every year), so I don't need to change that, but I would change my mean value of alpha (the intercept) to 120.
## 4M6
I would change the sigma estimate to 
sigma ~ uniform(0,8)
This would assume that the variance is always 64, which is conservative since it may at times be smaller.
I would also change the alpha estimate to 
alpha ~Normal(120,64) so that, even at time=0 the sigma value of 64 is observed.
# For 03/31/16

## 4H1
```{r}
library(rethinking) 
  data(Howell1) 
  d <- Howell1 
  str(d)
 model1 <- map( 
  alist(
  height ~ dnorm( mu , sigma ) , 
  mu <- a + b*weight , 
  a ~ dnorm( 178 , 100 ) , 
  b ~ dnorm( 0 , 10 ) , 
  sigma ~ dunif( 0 , 50 )
), 
data=d )
precis(model1, corr=TRUE) 
post <- extract.samples( m4.3 )
#46.95
  mu_at_46.95 <- post$a + post$b * 46.95
  dens( mu_at_46.95 , col=rangi2 , lwd=2 , xlab="mu|weight=46.95" )
  HPDI(mu_at_46.95, prob=0.89)
  weight.seq <- 46.95
  mu <- link( m4.3 , data=data.frame(weight=weight.seq) ) 
  str(mu)
  mu.mean <- apply( mu , 2 , mean ) 
  mu.mean
 #43.72 
 mu_at_43.72 <- post$a + post$b * 43.72
  dens( mu_at_43.72 , col=rangi2 , lwd=2 , xlab="mu|weight=43.72" )
  HPDI(mu_at_43.72, prob=0.89)
  weight.seq <- 43.72
  mu <- link( m4.3 , data=data.frame(weight=weight.seq) ) 
  str(mu)
  mu.mean <- apply( mu , 2 , mean ) 
  mu.mean
  #64.78 
 mu_at_64.78 <- post$a + post$b * 64.78
  dens( mu_at_64.78 , col=rangi2 , lwd=2 , xlab="mu|weight=64.78" )
  HPDI(mu_at_64.78, prob=0.89)
  weight.seq <- 64.78
  mu <- link( m4.3 , data=data.frame(weight=weight.seq) ) 
  str(mu)
  mu.mean <- apply( mu , 2 , mean ) 
  mu.mean
  
   #32.59 
 mu_at_32.59 <- post$a + post$b * 32.59
  dens( mu_at_32.59 , col=rangi2 , lwd=2 , xlab="mu|weight=32.59" )
  HPDI(mu_at_32.59, prob=0.89)
  weight.seq <- 32.59
  mu <- link( m4.3 , data=data.frame(weight=weight.seq) ) 
  str(mu)
  mu.mean <- apply( mu , 2 , mean ) 
  mu.mean
  
  #54.63 
 mu_at_54.63 <- post$a + post$b * 54.63
  dens( mu_at_54.63 , col=rangi2 , lwd=2 , xlab="mu|weight=54.63" )
  HPDI(mu_at_54.63, prob=0.89)
  weight.seq <- 54.63
  mu <- link( m4.3 , data=data.frame(weight=weight.seq) ) 
  str(mu)
  mu.mean <- apply( mu , 2 , mean ) 
  mu.mean
  
```
observation	weight	expected height	89% HPDI
1	46.95	158.2674	157.5-159.1
2	43.72	152.5759	151.86-153.33
3	64.78	189.6565	188.34-191.22
4	32.59	132.9472	132.27-133.57
5	54.63	171.8163	170.67-172.88
## 4H2
#a
```{r}
library(rethinking) 
  data(Howell1) 
  d2 <- d[ d$age < 18 , ] 
  str(d2)
  m4H2 <- map( 
  alist(
  height ~ dnorm( mu , sigma ) , 
  mu <- a + b*weight , 
  a ~ dnorm( 138 , 100 ) , 
  b ~ dnorm( 0 , 10 ) , 
  sigma ~ dunif( 0 , 50 )
), 
data=d2 )
precis(m4H2, corr=TRUE)
```
With a slope of 2.72, for every 10 unit increase in weight, the height should increase by 27.2
#b
```{r}
library(rethinking)
post2 <- extract.samples(m4H2)
mu.link <- function(weight) post2$a + post2$b*weight
weight.seq<- seq(0,40, by=1)
mu <- sapply(weight.seq, mu.link)
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob=0.89)
plot(d2$height ~ d2$weight, xlab="weight", ylab="height")
lines(weight.seq, mu.mean)
shade(mu.HPDI, weight.seq)
pred.height <- sim(m4H2, data=list(weight=weight.seq))
height.PI <- apply(pred.height, 2, PI, prob=0.89)
shade(height.PI, weight.seq)
```
The model is appropriate for linear data it seems, but the actual data points are not really showing a linear relationship. y increases with x only up to x of about 30, and then y levels off as x increases. 

  
## 4H3
```{r}
data <- Howell1
str(data)
data$logweight <- log(data$weight + 1) 
model4H3 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * logweight,
    a~ dnorm(178, 100),
    b ~ dnorm(0, 100),
    sigma ~ dunif(0, 50)
  ),
data = data)

plot(data$height, data$logweight)
```

```{r}
post3 <- extract.samples(model4H3)
mu.link <- function(weight) post3$a + post3$b*log(weight + 1)
weight.seq <- seq(0, 70, by=1)
mu <- sapply(weight.seq, mu.link)
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob=0.89)
lines(weight.seq, mu.mean)
shade(mu.HPDI, weight.seq)
pred_height <- sim(model4H3, data=list(weight=weight.seq))
height.PI <- apply(pred_height, 2, PI, prob=0.89)
shade(height.PI, weight.seq)
```